---
title: "Final Project: AirBnb Predictions"
author: "Julie Kim"
date: "04/30/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(pander)
library(dplyr)
library(pander)
library(geosphere)
library(glmnet)
library(randomForest)
library(stringr)
library(e1071)
library(rpart)
library(ada)
```


###Introduction

This project concerns various prediction tasks on AirBnB data from listings in 
Seattle (http://insideairbnb.com/get-the-data.html). We are interested in 
understanding the dataset better through the lens of classification, regression,
and clustering algorithms.

It is acceptable if the write-up is divided into 2 sections, Classification and 
Regression. And then within that, there should be individual introduction, 
exploration, supervised analysis, and results subsections.


###Task 1: Price Prediction

We would like to predict the Price for listings. A natural use case for this 
regressor would be in helping people price new listings. This task uses  
file `price.csv`.

1. Data exploration and pre-processing.
    ```{r}
    price = read.csv("price.csv")
    # is the data free of NA's
    all(!is.na(price))
    dim(price)
    
    # convert `host_is_superhost` to discrete numeric
      # f: 0
      # t: 1
    price$host_is_superhost = ifelse(price$host_is_superhost == "t", 1, 0)
    
    # convert `host_response_rate` to numeric in (0, 1)
    price$host_response_rate = gsub("%", "", as.character(price$host_response_rate)) %>% 
      as.numeric()/100
    
    # convert `host_response_time` to discrete numeric
      # within an hour: 0
      # within a few hours: 1
      # within a day: 3
      # a few days or more: 4
    num = ifelse(price$host_response_time == "within an hour", 0, 1)
    num[price$host_response_time == "within a day"] = 2
    num[price$host_response_time == "a few days or more"] = 3
    price$host_response_time = num
    
    # convert `host_identity_verified` to discrete numeric
      # f: 0
      # t: 1
    price$host_identity_verified = ifelse(price$host_identity_verified == "t", 1, 0)
    
    # convert `latitude` and `longitude` to distance to the landmarks below
    # distance in meters
      # Downtown Seattle (Pike Place Market)
      # Seattle-Tacoma International Airport
    # omit `latitude`, `longitude`, and `neighbourhood_group_cleansed`
    SEA.lat = 47.4502499
    SEA.lon = -122.3438354
    PP.lat = 47.6097199
    PP.lon = -122.3465703
    SEA = numeric()
    PP = numeric()
    for (i in 1:nrow(price)) {
      SEA[i] = distm(c(price$longitude[i], price$latitude[i]), 
                     c(SEA.lon, SEA.lat), fun = distHaversine)
      PP[i] = distm(c(price$longitude[i], price$latitude[i]),
                     c(PP.lon, PP.lat), fun = distHaversine)
    }
    price$Airport = SEA
    price$Downtown = PP
    price = price[,!(colnames(price) %in% c("latitude", "longitude", 
                                           "neighbourhood_group_cleansed"))]
    
    # convert `property_type` to discrete numeric
      # Apartment: 0
      # House: 1
      # Serviced: 2
      # Rustic: 3
      # Other: 4
    num = ifelse(price$property_type %in% c("Apartment", "Condominium", "Loft"), 
                 0, 3)
    num[price$property_type %in% c("House", "Townhouse", "Guesthouse",
                                   "Bungalow", "Cottage", "Cabin", "Villa")] = 1
    num[price$property_type %in% c("Guest suite", "Serviced apartment", 
                                   "Aparthotel", "Bed and breakfast",
                                   "Boutique hotel", "Resort", "Hostel")] = 2
    price$property_type = num
    
    # convert `room_type` to discrete numeric
      # Entire home/apt: 0
      # Private room: 1
      # Shared room: 2
    num = ifelse(price$room_type == "Entire home/apt", 0, 1)
    num[price$room_type == "Shared room"] = 2
    price$room_type = num
    
    # convert `bed_type` to discrete numeric
      # Real Bed: 0
      # Futon: 1
      # Pull-out Sofa: 2
      # Airbed: 3
      # Couch: 4
    num = ifelse(price$bed_type == "Real Bed", 0, 1)
    num[price$bed_type == "Pull-out Sofa"] = 2
    num[price$bed_type == "Airbed"] = 3
    num[price$bed_type == "Couch"] = 4
    price$bed_type = num
    
    # convert `amenities` to number of amenities
    price$amenities = strsplit(as.character(price$amenities), ",") %>%
      sapply(length)
    
    # convert `cleaning fee` to numeric
    split = ifelse(price$cleaning_fee == 0, 0, 
                   strsplit(as.character(price$cleaning_fee), c("\\$")) %>%
                     sapply(function(x) return(x[2])))
    num = as.numeric(gsub(",", "", split))
    price$cleaning_fee = num
    
    # convert `instant_bookable` to discrete numeric
      # f: 0
      # t: 1
    price$instant_bookable = ifelse(price$instant_bookable == "t", 1, 0)
    
    # convert `cancellation_policy` to discrete numeric
      # strict_14_with_grace_period: 0
      # moderate: 1
      # strict: 2
      # flexible: 3
      # super_strict_30: 4
      # super_strict_60: 5
    num = ifelse(price$cancellation_policy == "strict_14_with_grace_period", 0, 1)
    num[price$cancellation_policy == "strict"] = 2
    num[price$cancellation_policy == "flexible"] = 3
    num[price$cancellation_policy == "super_strict_30"] = 4
    num[price$cancellation_policy == "super_strict_60"] = 5
    price$cancellation_policy = num
    
    # convert `price` to numeric
    split = strsplit(as.character(price$price), c("\\$")) %>%
      sapply(function(x) return(x[2]))
    num = as.numeric(gsub(",", "", split))
    price$price = num
    
    # omit `X` and `id`
    price = price[,!(colnames(price) %in% c("X", "id"))]
    dim(price)
    ```

2. Building and validation of predictive algorithms. 
    
    Split
    ```{r}
    set.seed(100)
    # split `price` into training and validation sets, 75/25
    i = sample(1:nrow(price), nrow(price)*0.25)
    train = price[-i,]
    validation = price[i,]
    
    # construct matrix x and response y for model fit
    x = model.matrix(price ~ . - 1, train)
    y = train$price
    x.val = model.matrix(price ~ . - 1, validation)
    y.val = validation$price
    ```
    
    Logistic Regression
    ```{r}
    set.seed(100)
    lg.mdl = glm(price ~ ., data = train, family = gaussian)
    # predictions on the validation data
    lg.pred.t = predict(lg.mdl, train)
    lg.pred.v = predict(lg.mdl, validation)
    # MSE
    lg.mse.t = mean((y-lg.pred.t)^2)
    lg.mse.v = mean((y.val-lg.pred.v)^2)
    # predictions vs. observations
    plot(y.val, lg.pred.v, 
         main = "Model Predictions vs. True Observations",
         xlab = "Price", ylab = "Predictions")
    ```
    
    Shrinkage Methods
    ```{r, autodep = TRUE, cache = TRUE}
    set.seed(100)
    ### Lasso ###
    lasso.mdl = cv.glmnet(x, y, family = "gaussian", alpha = 1)
    # minimizing lambda
    lasso.mdl$lambda.min
    # model coefficients
    # coef(lasso.mdl)
    # predictions on the validation data
    lasso.pred.t = predict(lasso.mdl, x)
    lasso.pred.v = predict(lasso.mdl, x.val)
    # MSE
    lasso.mse.t = mean((y-lasso.pred.t)^2)
    lasso.mse.v = mean((y.val-lasso.pred.v)^2)
    # predictions vs. observations
    plot(y.val, lasso.pred.v, 
         main = "Model Predictions vs. True Observations",
         xlab = "Price", ylab = "Predictions")
    
    ### Ridge ###
    ridge.mdl = cv.glmnet(x, y, family = "gaussian", alpha = 0)
    # minimizing lambda
    ridge.mdl$lambda.min
    # model coefficients
    # coef(ridge.mdl)
    # predictions on the validation data
    ridge.pred.t = predict(ridge.mdl, x)
    ridge.pred.v = predict(ridge.mdl, x.val)
    # MSE
    ridge.mse.t = mean((y-ridge.pred.t)^2)
    ridge.mse.v = mean((y.val-ridge.pred.v)^2)
    # predictions vs. observations
    plot(y.val, ridge.pred.v, 
         main = "Model Predictions vs. True Observations",
         xlab = "Price", ylab = "Predictions")
    
    ### Elastic Net ###
    en.mdl = cv.glmnet(x, y, family = "gaussian", alpha = 0.5)
    # minimizing lambda
    en.mdl$lambda.min
    # model coefficients
    # coef(ridge.mdl)
    # predictions on the validation data
    en.pred.t = predict(en.mdl, x)
    en.pred.v = predict(en.mdl, x.val)
    # MSE
    en.mse.t = mean((y-en.pred.t)^2)
    en.mse.v = mean((y.val-en.pred.v)^2)
    # predictions vs. observations
    plot(y.val, en.pred.v, 
         main = "Model Predictions vs. True Observations",
         xlab = "Price", ylab = "Predictions")
    ```
    
    Random Forest
    ```{r}
    set.seed(100)
    #diagnosis for second full dataset
    rf.mdl = randomForest(x, y, ntree = 400)
    # predictions on the validation data
    rf.pred.t = predict(rf.mdl, x)
    rf.pred.v = predict(rf.mdl, x.val)
    # MSE
    rf.mse.t = mean((y-rf.pred.t)^2)
    rf.mse.v = mean((y.val-rf.pred.v)^2)
    # predictions vs. observations
    plot(y.val, rf.pred.v, 
         main = "Model Predictions vs. True Observations",
         xlab = "Price", ylab = "Predictions")
    # error vs number of trees
    plot(rf.mdl)
    # variable importance plot
    varImpPlot(rf.mdl, sort = T, n.var = 10, 
               main = "Top 10 - Variable Importance")
    ```
    
    Comparison of MSE's
    ```{r}
    data.frame(model = c("Logistic Regression", "Lasso Regression", 
                         "Ridge Regression", "Elastic Net Regression", 
                         "Random Forest"), 
               train_MSE = c(lg.mse.t, lasso.mse.t, ridge.mse.t, en.mse.t, 
                             rf.mse.t), 
               validation_MSE = c(lg.mse.v, lasso.mse.v, ridge.mse.v, en.mse.v, 
                                  rf.mse.v))
    ```

    Pre-processing of the test data
    ```{r}
    test = read.csv("price_test.csv")
    # is the data free of NA's
    all(!is.na(test))
    dim(test)
    
    # convert `host_is_superhost` to discrete numeric
      # f: 0
      # t: 1
    test$host_is_superhost = ifelse(test$host_is_superhost == "t", 1, 0)
    
    # convert `host_response_rate` to numeric in (0, 1)
    test$host_response_rate = gsub("%", "", as.character(test$host_response_rate)) %>% 
      as.numeric()/100
    
    # convert `host_response_time` to discrete numeric
      # within an hour: 0
      # within a few hours: 1
      # within a day: 3
      # a few days or more: 4
    num = ifelse(test$host_response_time == "within an hour", 0, 1)
    num[test$host_response_time == "within a day"] = 2
    num[test$host_response_time == "a few days or more"] = 3
    test$host_response_time = num
    
    # convert `host_identity_verified` to discrete numeric
      # f: 0
      # t: 1
    test$host_identity_verified = ifelse(test$host_identity_verified == "t", 1, 0)
    
    # convert `latitude` and `longitude` to distance to the landmarks below
    # distance in meters
      # Downtown Seattle (Pike Place Market)
      # Seattle-Tacoma International Airport
    # omit `latitude`, `longitude`, and `neighbourhood_group_cleansed`
    SEA.lat = 47.4502499
    SEA.lon = -122.3438354
    PP.lat = 47.6097199
    PP.lon = -122.3465703
    SEA = numeric()
    PP = numeric()
    for (i in 1:nrow(test)) {
      SEA[i] = distm(c(test$longitude[i], test$latitude[i]), 
                     c(SEA.lon, SEA.lat), fun = distHaversine)
      PP[i] = distm(c(test$longitude[i], test$latitude[i]),
                     c(PP.lon, PP.lat), fun = distHaversine)
    }
    test$Airport = SEA
    test$Downtown = PP
    test = test[,!(colnames(test) %in% c("latitude", "longitude", 
                                           "neighbourhood_group_cleansed"))]
    
    # convert `property_type` to discrete numeric
      # Apartment: 0
      # House: 1
      # Serviced: 2
      # Rustic: 3
      # Other: 4
    num = ifelse(test$property_type %in% c("Apartment", "Condominium", "Loft"), 
                 0, 3)
    num[test$property_type %in% c("House", "Townhouse", "Guesthouse",
                                   "Bungalow", "Cottage", "Cabin", "Villa")] = 1
    num[test$property_type %in% c("Guest suite", "Serviced apartment", 
                                   "Aparthotel", "Bed and breakfast",
                                   "Boutique hotel", "Resort", "Hostel")] = 2
    test$property_type = num
    
    # convert `room_type` to discrete numeric
      # Entire home/apt: 0
      # Private room: 1
      # Shared room: 2
    num = ifelse(test$room_type == "Entire home/apt", 0, 1)
    num[test$room_type == "Shared room"] = 2
    test$room_type = num
    
    # convert `bed_type` to discrete numeric
      # Real Bed: 0
      # Futon: 1
      # Pull-out Sofa: 2
      # Airbed: 3
      # Couch: 4
    num = ifelse(test$bed_type == "Real Bed", 0, 1)
    num[test$bed_type == "Pull-out Sofa"] = 2
    num[test$bed_type == "Airbed"] = 3
    num[test$bed_type == "Couch"] = 4
    test$bed_type = num
    
    # convert `amenities` to number of amenities
    test$amenities = strsplit(as.character(test$amenities), ",") %>%
      sapply(length)
    
    # convert `cleaning fee` to numeric
    split = ifelse(test$cleaning_fee == 0, 0, 
                   strsplit(as.character(test$cleaning_fee), c("\\$")) %>%
                     sapply(function(x) return(x[2])))
    num = as.numeric(gsub(",", "", split))
    test$cleaning_fee = num
    
    # convert `instant_bookable` to discrete numeric
      # f: 0
      # t: 1
    test$instant_bookable = ifelse(test$instant_bookable == "t", 1, 0)
    
    # convert `cancellation_policy` to discrete numeric
      # strict_14_with_grace_period: 0
      # moderate: 1
      # strict: 2
      # flexible: 3
      # super_strict_30: 4
      # super_strict_60: 5
    num = ifelse(test$cancellation_policy == "strict_14_with_grace_period", 0, 1)
    num[test$cancellation_policy == "strict"] = 2
    num[test$cancellation_policy == "flexible"] = 3
    num[test$cancellation_policy == "super_strict_30"] = 4
    num[test$cancellation_policy == "super_strict_60"] = 5
    test$cancellation_policy = num
    
    # omit `X` and `id`
    test = test[,!(colnames(test) %in% c("X", "id"))]
    dim(test)
    ```
    
    `price.guesses`: A single vector of price estimates
    ```{r}
    set.seed(100)
    # construct matrix x and response y for test predictions
    x.test = model.matrix(~ . - 1, test)
    # predictions on the validation data
    price.guesses = round(predict(rf.mdl, x.test), digits = 2)
    ```
   
    `price.mse`: A single number indicating your best guess at the mean squared error for your predictions on the test set
    ```{r}
    price.mse = 6300
    ```
    
    `review.guesses`: A single binary {0,1} vector of predictions of whether the test listings are undesirable or desirable. This vector should be as long as the number of test cases in review binary test.csv (and in the same order).

    `review.acc`: A single number indicating your best guess at the 0/1 error for your predictions on the test set.
    
    `team.name`: A string with your team’s name. These will be revealed in class, so make it anonymous if you wish. Your report will link the team name to individual names for grading purposes.
    ```{r}
    team.name = "boolin"
    ```
   
3. Actual submission to a prediction contest

4. Some follow-up analysis of your results. 

5. Two (optional) extra-credit exercises.


###Task 2: Review Classifier

Each AirBnB listing also has a review score. I have discretized this variable 
into two classes {0, 1} corresponding to {undesirable, desirable}. Our goal will 
be to design a classifier for this variable. Use the file `review.csv`.

```{r}
review <- read.csv("review.csv")
```

#1 Cleaning & Splitting Data

## Cleaning Data & Modifying Variables
Here I write a function to clean my data which modifies/combines/creates/eliminates variables from the original dataset and returns a new dataset with all my desired predictors.

```{r, predictors generator function}
predictors <- function(r){
  
  #host variables
  responserate <-as.numeric(str_remove(r$host_response_rate, "%"))/100
  responsetimelessday <- as.numeric(ifelse(r$host_response_time == "a few days or more", 0, 1))
  superhost <- as.numeric(r$host_is_superhost)-1
  verified <- as.numeric(r$host_identity_verified)-1
  listings <- as.numeric(r$host_listings_count)
  
  #about the place
  #amenities conversions
  n_amenities <- str_count(r$amenities, ",")
  tv <-ifelse(str_count(tolower(r$amenities), "tv") == 0, 0, 1)
  wifi <-ifelse(str_count(tolower(r$amenities), "wifi") == 0, 0, 1)
  kitchen <-ifelse(str_count(tolower(r$amenities), "kitchen") == 0, 0, 1)
  heating <-ifelse(str_count(tolower(r$amenities), "heating") == 0, 0, 1)
  refrigerator <-ifelse(str_count(tolower(r$amenities), "refrigerator") == 0, 0, 1)
  microwave <-ifelse(str_count(tolower(r$amenities), "microwave") == 0, 0, 1)
  parking <-ifelse(str_count(tolower(r$amenities), "parking") == 0, 0, 1)
  washer <-ifelse(str_count(tolower(r$amenities), "washer") == 0, 0, 1)
  dryer <-ifelse(str_count(tolower(r$amenities), "dryer") == 0, 0, 1)
  essentials <-ifelse(str_count(tolower(r$amenities), "essentials") == 0, 0, 1)
  shampoo <-ifelse(str_count(tolower(r$amenities), "shampoo") == 0, 0, 1)
  stove <-ifelse(str_count(tolower(r$amenities), "stove") == 0, 0, 1)
  #distance to aquarium - as central/tourist downtown seattle location
  longlats <- r[,c(12, 11)]
  distaquarium <- as.numeric(as.vector(distm(longlats, c(-122.342463, 47.607515))))
  #description of place
  typeproperty <- ifelse(r$property_type == "Aparthotel", "serviced", "")
  typeproperty[which(r$property_type %in% c("Serviced apartment", "Bed and breakfast", "Boutique hotel"))] <- "serviced"
  typeproperty[which(r$property_type %in% c("Cottage", "Bungalow", "Cabin","Guesthouse", "House", "Townhouse", "Villa"))] <- "standalone"
  typeproperty[which(r$property_type %in% c("Yurt", "Tiny house", "Other", "Houseboat", "Boat", "Camper/RV", "Tent", "Farm stay"))] <- "weird"
  typeproperty[which(r$property_type %in% c("Loft", "Condominium", "Apartment"))] <- "complex"
  typeproperty[which(r$property_type %in% c("Guest suite", "Hostel", "In-law"))] <- "share"
  typeroom <- r$room_type
  fits <- as.numeric(r$accommodates)
  baths <- as.numeric(r$bathrooms)
  bedrooms <- as.numeric(r$bedrooms)
  beds <- as.numeric(r$beds)
  realbed <- ifelse(r$bed_type == "Real Bed", 1, 0)
  #fees and pricing
  split = strsplit(as.character(r$price), c("\\$")) %>%
    sapply(function(x) return(x[2]))
  price <- as.numeric(gsub(",", "", split))
  #clean 0 cleaning fee to $0 cleaning fee
  r[which(r$cleaning_fee == 0), "cleaning_fee"] = "$0.00"
  split = strsplit(as.character(r$cleaning_fee), c("\\$")) %>%
    sapply(function(x) return(x[2]))
  cleanfee <- as.numeric(gsub(",", "", split))
  #logistics
  minnights <- as.numeric(r$minimum_nights)
  maxnights <- as.numeric(r$maximum_nights)
  strictcancelpol <- ifelse(r$cancellation_policy == "flexible", 0, 1)
  strictcancelpol[r$cancellation_policy == "moderate"] <- 0
  #put predictors in dataframe and return
  df <- data.frame(responserate = responserate, responsetimelessday = responsetimelessday, superhost = superhost, verified = verified, listings = listings, n_amenities = n_amenities, tv = tv, wifi = wifi, kitchen = kitchen, heating = heating, refrigerator = refrigerator, microwave = microwave, parking = parking, washer = washer, dryer = dryer, essentials = essentials, shampoo = shampoo, stove = stove, distaquarium = distaquarium, typeproperty = typeproperty, typeroom = typeroom, fits = fits, baths = baths, bedrooms = bedrooms, beds = beds, realbed = realbed, price = price, cleanfee = cleanfee, minnights = minnights, maxnights = maxnights, strictcancelpol = strictcancelpol)
  return(df)
}
```

## Splitting the Data

Shuffling the data and splitting the data into 75% training, and 25% testing. Also reporting the rate of naive prediction, that is guessing the most common class.

```{r, binding score and splitting data into train & test}
preds <- predictors(review)
score <- data.frame(score = review$review_scores_rating)
review.new <- cbind.data.frame(score, preds)
set.seed(1000)
shuffle <- review.new[sample.int(nrow(review.new),nrow(review.new)),]
train <- shuffle[1:3032,]
test <- shuffle[3033:nrow(shuffle),]
#naive classification, pick most common class
misclas.naive <- length(which(test$score == 0))/length(test$score)
misclas.naive
```


#Model Selection

I am testing a variety of models on the training set and returning misclassification error on the test set.

##Logistic Binomial

```{r, logistic model}
#logistic model
log1 <- glm(score ~., data = train, family = "binomial")
pred.log1 <- ifelse(predict(log1, newdata = test, type = "response") < .5, 0, 1)
misclass.log1 <- sum(as.numeric(pred.log1 != test$score), na.rm = T)/length(na.omit(pred.log1))
misclass.log1
```


##Support Vector Machines

Ran with 5 kernels: linear, radial, polynomial with k = {2,3,4}

```{r, svms}
#support vector machines 
#linear boundary
svm.lin <- svm(score ~., type = "C-classification", data = train, kernel = "linear")
pred.svm.lin <- predict(svm.lin, newdata = test)
misclas.svm.lin <- sum(as.numeric(pred.svm.lin != na.omit(test)$score))/length(pred.svm.lin)
misclas.svm.lin
#radial boundary
svm.rad <- svm(score ~., type = "C-classification", data = train, kernel = "radial")
pred.svm.rad <- predict(svm.rad, newdata = test)
misclas.svm.rad <- sum(as.numeric(pred.svm.rad != na.omit(test)$score))/length(pred.svm.rad)
misclas.svm.rad
#k = 2 polynomial boundary
svm.pol2 <- svm(score ~., type = "C-classification", data = train, kernel = "polynomial", degree = 2)
pred.svm.pol2 <- predict(svm.pol2, newdata = test)
misclas.svm.pol2 <- sum(as.numeric(pred.svm.pol2 != na.omit(test)$score))/length(pred.svm.pol2)
misclas.svm.pol2
#k = 3 polynomial boundary
svm.pol3 <- svm(score ~., type = "C-classification", data = train, kernel = "polynomial", degree = 3)
pred.svm.pol3 <- predict(svm.pol3, newdata = test)
misclas.svm.pol3 <- sum(as.numeric(pred.svm.pol3 != na.omit(test)$score))/length(pred.svm.pol3)
misclas.svm.pol3
#k = 4 polynomial boundary
svm.pol4 <- svm(score ~., type = "C-classification", data = train, kernel = "polynomial", degree = 4)
pred.svm.pol4 <- predict(svm.pol4, newdata = test)
misclas.svm.pol4 <- sum(as.numeric(pred.svm.pol4 != na.omit(test)$score))/length(pred.svm.pol4)
misclas.svm.pol4
```

##Decision Trees

Reporting error on full and pruned tree.

```{r, trees}
#basic tree
tree1 <- rpart(score ~ ., data = train)
pred.tree1 <- ifelse(predict(tree1, newdata = test) > .5, 1, 0)
misclas.tree1 <- sum(pred.tree1 != test$score)/length(pred.tree1)
misclas.tree1
#pruned tree
#plotcp(tree1) #used to determine where to prune
tree2 <- prune(tree1, 0.014)
pred.tree2 <- ifelse(predict(tree2, newdata = test) > .5, 1, 0)
misclas.tree2 <- sum(pred.tree2 != test$score)/length(pred.tree2)
misclas.tree2 
```

##Random Forests

Did random forest with score as a factor, then with score as numeric and bounded the responses.

```{r, random forests, warning = F, message = F}
#regular random forest score as factor
set.seed(1000)
rf1.factor <- randomForest(factor(score) ~., data = na.omit(train))
pred.rf1.factor <- predict(rf1.factor, newdata = na.omit(test))
misclas.rf1.f <- sum(pred.rf1.factor != na.omit(test)$score)/length(pred.rf1.factor)
misclas.rf1.f
#score not as factor, predict as bounded probabilities
set.seed(1000)
rf1 <- randomForest(score ~., data = train)
pred.rf1 <- ifelse(predict(rf1, newdata = test) > .5, 1, 0)
misclas.rf1 <- sum(pred.rf1 != test$score)/length(pred.rf1)
misclas.rf1
```

##Adaptive Boosting

Performed boosting first with exponential loss, then with logistic loss.

```{r, AdaBoost}
#exponential loss
set.seed(1000)
ada1 <- ada(score ~., data = train, loss = "exponential")
pred.ada1 <- predict(ada1, newdata = test, type = "vector")
misclas.ada1 <- sum(pred.ada1 != test$score)/length(test$score)
misclas.ada1
#logistic loss
set.seed(1000)
ada2 <- ada(score ~., data = train, loss = "logistic")
pred.ada2 <- predict(ada2, newdata = test, type = "vector")
misclas.ada2 <- sum(pred.ada2 != test$score)/length(test$score)
misclas.ada2 #(smaller than exponential)


```

##Model Selection

I am selecting the non-factor random forest.

#Prediction

Reading in data, predicting. Best guess at error is misclassification rate of model on test set.

```{r, generating review.guesses and review.acc}
#generating predictions with the balanced random forest model
review.test <- read.csv("review_test.csv")
test.dat <- predictors(review.test)
review.guesses <-predict(ada1, newdata = test.dat)
review.acc <- misclas.ada1
```

###Making predictions

How can you make your predictions? You can use any of the techniques we have 
discussed in class. You can use any of the variables in the data set, and you 
can also consider constructing new variables by combining or transforming the 
variables that are present in the data set. You should not use external 
information sources for this project.

###Submitting predictions

You will submit a single RData file with your predictions. This file should 
contain the following variables:
1. price.guesses: A single vector of price estimates. This vector should be as 
   long as the number of test cases in price_test.csv (and in the same order).
2. price.mse: A single number indicating your best guess at the mean squared 
   error for your predictions on the test set. This will give us an idea of how 
   well your validation has worked in setting your expectations. Good estimation 
   of your performance is one goal of the project.
3. review.guesses: A single binary {0,1} vector of predictions of whether the 
   test listings are undesirable or desirable. This vector should be as long as 
   the number of test cases in review binary test.csv (and in the same order).
4. review.acc: A single number indicating your best guess at the 0/1 error for 
   your predictions on the test set.
5. team.name: A string with your team’s name. These will be revealed in class, 
   so make it anonymous if you wish. Your report will link the team name to 
   individual names for grading purposes.
   
To make this file, if you have the appropriate variables in your workspace, you 
can type `save(list=c("price.guesses","price.mse","review.guesses", "review.acc", "team.name"),file="stat462final.RData")`
This will create stat462final.RData file, which you can upload on canvas. 
(Please rename the file to include your team name before sending.)

